### 2013-11-23 paper/interview1.tex
Added start of interview.

\clearpage

\section*{Appendix A: Interview Transcript}
Interview with Traci St. Martin, PMP\textregistered, Mayo Clinic, Nov 21, 2013.

\subsection*{Q1: How do development teams collaborate and share ideas?}

\subsection*{Q2: Do teams generally remain intact over multiple projects, or do members often shuffle and re-combine?}

\subsection*{Q3: When assembling a team, which is preferred -- members familiar with each other, or members familiar with the task or project.}

\subsection*{Q4: What are some characteristics of a high performing team?}

\subsection*{Q5: In your experience, is it better to (A) keep an innovative team intact, (B) periodically bring in new members or re-assign existing members, or (C) split up and disperse the team to as many other projects as possible?}