### 2013-11-30 paper/main.tex
Main content in... the revisions begin.

\documentclass{proc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{fixltx2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{tikz}
\usetikzlibrary{positioning}
\setlength{\abovecaptionskip}{2pt}
\setcounter{secnumdepth}{2}

\title{
The Small World Network Effect in Software Project Teams
\author{Kevin Peterson\\
\small \texttt{pete1968@umn.edu}
}
}

\begin{document}
\maketitle

\begin{abstract}
Team cohesion and the dynamics of team forming are important parts of any project, with software projects being no exception. An interesting aspect of team building is the relationships formed between the team members. If follows that visualizing software team members as a graph can give some insight as to what the optimal team conditions are. As team members move between projects, these graphs become more and more connected as team members collaborate and form relationships. We show that this connectivity, known as the ``small world effect," has a positive impact on team performance when the connectivity levels are moderate. Performance degrades, however, at both very high and very low levels of connectivity. This aligns with similar research findings of non-software teams.
\end{abstract}

\noindent \\\textbf{Keywords.} Small World, Project Management, Software

\section{Introduction}
A social network is a graph of people and the connections between them. The dynamics of social networks have proved to be an important research topic for many different areas of study, from academic publication\cite{barabasi2002evolution}, to the success of broadway musicals\cite{uzzi2005collaboration}. In general, a social network is important to understanding how ideas and influence are spread\cite{kempe2003maximizing}, from which we can investigate the best conditions for optimal group perfomance within these networks.

A `small world' network is a social network characterized by high clustering of graph nodes paired with a short average path between nodes\cite{watts1998collective}. `Clustering' in this context means how closely nodes in the graph are related to each other, or "the friend of my friend is also my friend." Milgram, in is seminal study, found that any two people are linked through a chain of friends and acquaintances on average of 6 people long\cite{milgram1967small}, or in other words, the average `path length' between any two people.

Uzzi \cite{uzzi2005collaboration} expanded on this by exploring how these small world networks correlate to collaboration and creativity. In his study, the network of artists involved in the making of Broadway musicals from 1945 to 1989 were studied. Performance of the musicals were then measured based on financial success and critical acclaim, and compared that quantitative analysis to the clustering and path length of the graph.

We follow Uzzi's work, and expand it to software teams. We hypothesize that the qualities of a small world graph that foster collaboration, creativity, and the efficiency of how knowledge is transfered\cite{latora2001efficient} will also apply to software project teams. In this work, we investigate the `small world' phenomenon quantitatively by exploring open source project data, and qualitatively by a series of interviews of subject matter experts, and look to compare Uzzi's findings to ours.

\subsection{Hypothesis}
The formalization of the `small world' quality of a graph can be expressed by the \textit{Small World Quotient}, or $Q$\cite{watts1999small,watts1998collective}. The value $Q$ is calculated by dividing the clustering factor of graph by the average path length between nodes. Uzzi explored this value in graphs of Broadway productions\cite{uzzi2005collaboration}. His findings suggest that the success of Broadway productions is greatest when teams have a good mix of new and familiar members. Productions are less likely to be successful when team members are very familiar with each other. Also, probability of success drops much the same way when there is little familiarity between team members. This indicates that there is an optimal ratio of new and familiar members on a team -- or, an optimal value of $Q$ when related to performance.

\textit{Software Project Teams will perform best at moderate levels of $Q$. This performance, measured by the amount of interest their projects generate, will increase as the level of $Q$ increases for the contributor graph. This will continue up to an optimal $Q$ value, after which performance will begin to degrade. Thus, the performance curve given $Q$ will be inverse U-shaped, and match Uzzi's findings\cite{uzzi2005collaboration}.}

\section{Methods}

\subsection{Collection and Storage}
Data was collected using the freely available FLOSSmole\cite{floss2006} data sets of free and open source projects. Two datasets were analyzed -- data from from the popular open source project site \textit{Freecode},\footnote{http://freecode.com/}, and another from open source software repository \textit{SourceForge}. The Freecode dataset contained data up to September, 2013, while the SourceForge dataset was slightly older with data gathered on June 2009. Both datasets were chosen because they met two basic criteria: (1) they contained an exhaustive list of projects and contributors to those projects, and (2) they provided some sort of quantitative measure of project ``popularity." This popularity metric will be further described in sectiions below. Results were analyzed using the Python graph processing packaged NetworkX\cite{hagberg-2008-exploring}. All source code pertaining to the collection and analysis can be found at \url{https://github.com/kevinpeterson/small-world-effect-research}.

\subsection{Analysis}
The collected data was organized into a graph structure, with the nodes representing the collaborators and the edges representing one or more shared project between collaborators. Various properties of the graph are analyzed, followed by a qualitative section to help link the quantitative analysis to subject matter expert experiences.

\subsubsection{Graph Structure}
For each dataset, we first considered the dataset to be one large graph. From this large disconnected graph, each connected subgraph was analyzed. Metrics were then computed for each of these subgraphs. In order to correctly compute the metrics below, a minimum subgraph size was required. Specifically, many of the graph metrics of interest measured ``triangles'' of nodes and their connections. This implies a subgraph of {$>= 3$} nodes is needed. Subgraphs of {$< 3$} nodes were not considered for analysis.

\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{P/.style={circle,draw=black,fill=lightgray!20}}
\tikzset{C/.style={circle,draw=black,fill=black!20}}

 [scale=.5,auto=left]
  \node[P] (p1) at (-5,3)  {$P_1$};
  \node[P] (p2) at (-5,2)  {$P_2$};
  \node[P] (p3) at (-5,1)  {$P_3$};
  \node[P] (p4) at (-5,0)  {$P_4$};

  \node[C] (c1) at (-2,3)  {$C_1$};
  \node[C] (c2) at (-2,2)  {$C_2$};
  \node[C] (c3) at (-2,1)  {$C_3$};
  \node[C] (c4) at (-2,0)  {$C_4$};

  \draw ({p1}) -- ({c1});
  \draw ({p1}) -- ({c2});
  \draw ({p2}) -- ({c2});;
  \draw ({p3}) -- ({c2});
  \draw ({p3}) -- ({c3});
  \draw ({p4}) -- ({c3});
  \draw ({p4}) -- ({c4});
  \node[draw=none,fill=none,rectangle,left=1mm of p1] {Projects}; 
  \node[draw=none,fill=none,rectangle,right=1mm of c1] {Contributors}; 
\end{tikzpicture}
\caption{An example bipartite graph of Projects and Contributors}
\label{fig:example_bipartite_graph}
\end{figure}

The Freecode data can be represented as a bipartite graph. A bipartite graph is a specialized type of graph containing two disjoint sets of verticies. This can be denoted by ${G=(P,C,E)}$, where ${P \cap C = \emptyset}$. For our purposes, let set $P$ be the set of all projects, and set $C$ be the set of all contributors. Each project node, therefore, is only connected to contributor nodes, and vice versa, as show in figure \ref{fig:example_bipartite_graph}. This bipartite representation is typical of what is found in social and collaboration networks\cite{ramasco2004self}.

A bipartite graph projection is necessary to further analyze the graph. A bipartite projection involves taking the disjoint node sets $P$ and $C$, and representing the graph as relationships between only one of those sets of nodes. A projection onto the contributor nodes, or $C$, of figure \ref{fig:example_bipartite_graph} is show in figure \ref{fig:example_bipartite_projection_graph}. Here, contributors are connected directly, with edges occurring if a two contributors share one or more common projects. It is important to note that the projection could have been done in terms of the projects, not the contributors. This would have led to a graph with nodes of projects $P$, each being linked by sharing a common contributor. This approach is a common way of representing these bipartite graphs\cite{newman2001scientific}, and illustrates the fact that a bipartite projections are lossy compared to the data represented in the original graph\cite{zhou2007bipartite}.\\

One of the main sources of information loss is multiple connections between contributors, or contributors that share multiple projects. When doing a bipartite projection based on contributors $C$, a single link could represent one shared project or many. This is a source of data loss in the projection. The simple approach is to treat one connection the same as many connections between nodes. This is straightforward but lossy\cite{zhou2007bipartite,grossman1995portion}. A better approach is to assign a \textit{weight} to each edge representing repeated links -- or in our case -- multiple shared collaborators\cite{zha2001bipartite,barrat2004architecture}.

\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{C/.style={circle,draw=black,fill=black!20}}

 [scale=.5,auto=left]
  \node[C] (c1) at (-5,3)  {$C_1$};
  \node[C] (c2) at (-3,2)  {$C_2$};
  \node[C] (c3) at (-5,1)  {$C_3$};
  \node[C] (c4) at (-2,1)  {$C_4$};

  \draw ({c1}) -- ({c3});
  \draw ({c1}) -- ({c2});
  \draw ({c1}) -- ({c2});
  \draw ({c2}) -- ({c3});
  \draw ({c3}) -- ({c4});
\end{tikzpicture}
\caption{A contributor ($C$) projection of the figure~\ref{fig:example_bipartite_graph} bipartite graph}
\label{fig:example_bipartite_projection_graph}
\end{figure}

\subsubsection{Graph Analysis}
To support our hypotheses, there are several key metrics to be studied in regards to the aforementioned graph. Many of these metrics follow the work of Boccaletti \textit{et. al}\cite{boccaletti2006complex}, and other studies of similar graph types\cite{latora2001efficient,adamic1999small}.

\noindent\\\textit{\textbf{Clustering Coefficient - Transitivity ($C^\Delta$)}}\\
We calculate the \textit{Clustering Coefficient} (or $C^\Delta$) as a measure of the proportion of closed triangles in a graph\cite{newman2003structure}:
\[C^\Delta = 3 \times \frac{\text{number of triangles}}
                    {\text{number of connected triples}}\]

This metric shows us how closely related, or ``clustered,'' that the graph is. It is also the probability that two nodes will be connected if they share a common neighbor\cite{newman2003properties}. In our context, if contributors $C_1$ and $C_2$ both collaborate with a common contributor w $C_3$, this metric represents the probability that $C_1$ and $C_2$ will collaborate.

\noindent\\\textit{\textbf{Clustering Coefficient - Average ($\overline{C^{\lambda}}$)}}\\
Another approach to clustering is to take the average of each node's local clustering coefficient\cite{watts1998collective}.

Given our graph of nodes $P$ connected by edges $E$, or ${G=(P,E)}$, for any given node we can determine its \textit{neighborhood} ($N_i$), or the set of nodes directly connected to a given node.
\[ N_i = \{v_j : e_{ij} \in E \wedge e_{ji} \in E \wedge v_j \in P\} \]

Given this \textit{neighborhood}, we can then calculate how closely connected the nodes are. To do this, we take the number of actual connections between neighbors divided by the total possible number of connections, where $k_i$ represents the count of the neighbors of a node, or {$k_i = |N_i|$}
\[ C^{\lambda}_i = \frac{2|\{e_{jk}: v_j,v_k \in N_i, e_{jk} \in E\}|}{k_i(k_i-1)} \]

From this \textit{Local Clustering Coefficient} $C_i$, we can then take the average over the entire set of nodes in the graph.
\[ \overline{C^\lambda} = \frac{1}{|P|}\sum^{|P|}_{i=1}C^{\lambda}_i \]

It is important to note that although the terminology is similar\cite{uzzi2005collaboration}, $\overline{C^\lambda}$ and $C^\Delta$ are different measurements. Unless otherwise noted, the Transitivity version of this metric ($C^\Delta$) will be used for all further calculations and metrics.

\noindent\\\textit{\textbf{Average Path Length ($L$)}}\\
Also, calculating the \textit{Path Length} ($L$) allows us to determine how far removed developers in the graph are from one another. The path length is calculated as the average distance between any two node pairs in the graph:

\[L = \frac{1}{|P| \cdot (|P|-1)} \sum_{i \neq j}\lambda(v_i,v_j)\]

Where {$v_i \in P$}, {$v_j \in P$}, and {$\lambda(v_i,v_j)$} denotes the shortest path between the two nodes. We only consider path length for connected subgraphs, avoiding some complexities of this calculation on disconnected graphs\cite{boccaletti2006complex}.

\begin{figure}
\centering
\begin{tikzpicture}
  [scale=.5,auto=left,every node/.style={circle,draw=black,fill=gray!20}]
  \node (n1) at (15,8) {1};
  \node (n2) at (12,6)  {2};
  \node (n3) at (9,5)  {3};
  \node (n4) at (6,5)  {4};
  \node[draw=none,fill=none,rectangle,right=1mm of n1] {$C^\Delta = 0.0$};

  \draw ({n1}) -- ({n2});
  \draw ({n2}) -- ({n3});
  \draw ({n3}) -- ({n4});
\end{tikzpicture}

\begin{tikzpicture}
  [scale=.5,auto=left,every node/.style={circle,draw=black,fill=gray!20}]

  \node (n1) at (15,8) {1};
  \node (n2) at (12,5)  {2};
  \node (n3) at (9,5)  {3};
  \node (n4) at (11,8)  {4};
  \node[draw=none,fill=none,rectangle,right=1mm of n1] {$C^\Delta = 1.0$};

  \draw ({n1}) -- ({n2});
  \draw ({n2}) -- ({n3});
  \draw ({n3}) -- ({n4});
  \draw ({n1}) -- ({n4});
  \draw ({n1}) -- ({n3});
  \draw ({n2}) -- ({n4});
\end{tikzpicture}

\begin{tikzpicture}
  [scale=.5,auto=left,every node/.style={circle,draw=black,fill=gray!20}]
  \node (n1) at (15,7) {1};
  \node (n2) at (12,5)  {2};
  \node (n3) at (8,7)  {3};
  \node (n4) at (11,8)  {4};
  \node[draw=none,fill=none,rectangle,right=1mm of n1] {$C^\Delta = 0.6$};

  \draw ({n1}) -- ({n2});
  \draw ({n3}) -- ({n4});
  \draw ({n1}) -- ({n4});
  \draw ({n2}) -- ({n4});
\end{tikzpicture}
\caption{Example Clustering Coefficient ($C^\Delta$) values for sample graphs}
\label{fig:example_graphs}
\end{figure}

\noindent\\\textit{\textbf{Small World Quotient ($Q$)}}\\
The definition of a \textit{small world} graph can be formalized by the following equations\cite{humphries2008network,uzzi2005collaboration}.

First, let $\gamma^{\Delta}_g$ equal the ratio of the clustering coefficient of a given graph and a random graph.
\[\gamma^{\Delta}_g = \frac{C^{\Delta}_g }{C^{\Delta}_{rand} } \]

Next apply a similar pattern to the average path length $\lambda_g$.
\[\lambda_g = \frac{L_g }{L_{rand} } \]

Finally, the ratio of the above calculations yields the small world quotient, or $Q$, of the graph
\[Q = \frac{\gamma^{\Delta}_g }{\lambda_g} \]

We assume for throughout that any graph exhibiting {$Q > 1$} is a small-world graph.

\noindent\\\textit{\textbf{Performance}}\\
To further explore the data, a measurement of project performance was needed. As the explored datasets consisted entirely of open source projects, research indicates that \textit{quality}, \textit{use}, \textit{user satisfaction}, and \textit{impact} are possible metrics to measure performance or \textit{success}\cite{crowston2003defining}.\\

For both datasets, we base our performance metric on \textit{use}. For Freecode, we use the \textit{Popularity Score} metric, which is calculated the following formula\footnote{http://help.freecode.com/kb/statistics/how-do-you-measure-a-projects-popularity}:

\[ ((record\ hits + URL\ hits) \cdot (followers + 1))^{1/2} \]

For the SourceForge dataset, performance is based on the repository's internal \textit{rank} metric. Note that there was no attempt to normalize the two performance metrics between datasets. Because of this, no cross-dataset measurements can be made, as the performance metric is only valid in the context of the enclosing dataset.

\subsubsection{Qualitative Analysis}
So far we have explored, via data analysis, how closely project collaborators are connected to one another, and what correlations we can derive from this analysis. We have said little as to why these connections are made, how they are maintained, and what exactly a connection means at a personal level. Via qualitative analysis, we will attempt to enrich our data analysis by providing deeper context.\\
A connection between two people is much more than an edge between two nodes in a graph. It has been asserted that people tend to make connections with people with whom they share similarities\cite{mcpherson2001birds}. This idea of homophily, however, needs to be further expanded before we can make assertions about correlations to performance. It has been shown that demographic homophily is complex to measure, correlates little to performance outcomes, and in fact correlate negatively\cite{reagans2004make,lawrence1997perspective}. For our purposes, it is much more appropriate to examine how teams best leverage similarities of knowledge and information, as the flow of these assets can be a key driver to success\cite{nissen2002extended}.\\
In the same way we can explore the role of team familiarity. The two types of familiarity that we will explore are \textit{member} familiarity and \textit{task/project} familiarity\cite{harrison2003time}. These metrics, although not captured in data, could be important factors in \textit{how} teams form, and thus help us to understand \textit{why} teams perform the way that they do.\\

To further study these factors, we arranged interviews with two current professionals in the field. Both have had experience with project management and are certified Project Management Professionals (PMP)\textregistered. We proposed to them the following questions, in hopes of expanding further the questions posed above:

\noindent\\\textit{Q1: How do development teams collaborate and share ideas?}\\
\noindent\\\textit{Q2: Do teams generally remain intact over multiple projects, or do members often shuffle and re-combine?}\\
\noindent\\\textit{Q3: When assembling a team, which is preferred -- members familiar with each other, or members familiar with the task or project.}\\
\noindent\\\textit{Q4: What are some characteristics of a high performing team?}\\
\noindent\\\textit{Q5: In your experience, is it better to (A) keep an innovative team intact, (B) periodically bring in new members or re-assign existing members, or (C) split up and disperse the team to as many other projects as possible?}\\

\section{Results}

\subsection{Quantitative}

\input{tables/subgraphs_summary_freecode}
\input{tables/subgraphs_summary_sf}

Summary statistics for the graph analysis are captured in figure \ref{fig:summary_stats}.  Statistics analyzed included:\\
$Q$ - \textit{Small World Quotient}\\
$C^\Delta$ - \textit{Clustering Coefficient - Transitivity}\\
$\overline{C^\lambda}$ - \textit{Clustering Coefficient - Average}\\
$L$ - \textit{Average Path Length}\\
$N$ - \textit{Number of Nodes}\\
$E$ - \textit{Number of Edges}\\
As measured over the number of projects $P$, with the total number of contributors $C$. The number of connected subgraphs analyzed is shown as $P_c$.

A mean value of $Q$ was observed to be $> 1$ for both data sets, which meets the criteria for a small-world network\cite{humphries2008network}. The median, however, is 1.0 in both cases, which does not match the small world network criteria. A possible explanation of this is the power law distribution of $Q$ observed in both data sets (figures \ref{fig:q_fc_distribution} and \ref{fig:q_sf_distribution}. This indicates that a large proportion of $Q$ values are around 1.\\

$C^\Delta$ and $\overline{C^\lambda}$ both indicate values approaching 1. This observation indicates a high degree of clustering\cite{watts1998collective}. hat any indicates that the majority of subgraphs have a $Q$ and $C^\Delta$ value of close to either 0 or 1. Exploring this phenomenon further, we can derive that most graphs resemble the middle or the bottom graph in figure \ref{fig:example_graphs}, with relatively few resembling the top graph.\\

Note that $P_c$ values are relatively small compared to $P$ and $C$ values. It would be intuitive to assume that {$C \times P_c \approx P$} -- in other words, if we split the entire graph into connected subgraphs, the number of subgraphs multiplied by the average number of nodes (contributors) per subgraph should equal the total number of contributors ($C$). This, however, is not the case. In fact, $P_c$ is notably smaller than expected. Recall that during analysis, subgraphs of {$< 3$} nodes were not considered. We can assume then that a large number of subgraphs are very small, with only one or two nodes. Further analysis is needed to fully understand the implications of this.

To validate the hypothesis, we expect to find that intermediate values of $Q$ will be correlated with more popular projects. To explore this possibility, we first compare the popularity given $Q$ values as shown in figures \ref{fig:q_fc_graph} and \ref{fig:q_sf_graph}. The $Q$ values do display the U-shaped characteristic as compared to popularity, which aligns with the findings of Uzzi \textit{et. al}\cite{uzzi2005collaboration}.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-graph.png}
\end{center}
\caption{Freecode $Q$ and Project Popularity}
\label{fig:q_fc_graph}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/sf-graph.png}
\end{center}
\caption{SourceForge $Q$ and Project Popularity}
\label{fig:q_sf_graph}
\end{figure}

Distribution of $Q$ values are shown in figures \ref{fig:q_fc_distribution} and \ref{fig:q_sf_distribution}. The values of $Q$ generally follow a power-law distribution, with most values being at our around {$Q = 1$}. This indicates that many graphs are close to the delineation of being considered small-world\cite{humphries2008network}.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-q-histo.png}
\end{center}
\caption{Freecode $Q$ Distribution}
\label{fig:q_fc_distribution}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/sf-q-histo.png}
\end{center}
\caption{SourceForge $Q$ Distribution}
\label{fig:q_sf_distribution}
\end{figure}

Small-world-ness, or $Q$, is generally between 1 and 3 for most sub-graphs.

\subsection{Qualitative}
The research interviews provided a look into how some of these quantitative findings translate into actual industry practice. It also helped us to get a deeper understanding of what we had found in the data analysis, as we could begin to build some workplace context into our results. There were are few re-occurring themes from which we will base our analysis\cite{stmartin_interview,rana_interview}:

\noindent\\\textit{\textbf{New contributors are good, as long as the core team remains intact.}}\\
Both participants mentioned the idea of keeping a ``core" team intact. This was certainly an often cited notion in both interviews. This concept appears to align with the idea of a \textit{core/periphery}\cite{borgatti2000models}, where the graph under study contains one (or many) tightly interconnected subgraphs, with less tightly connected nodes radiating out from this core.

Although our data analysis does not include any examination of a core/periphery, it is an interesting direction for future research. Notably, has been suggested that individual performance in a group is highest if that individual is not in the core or the periphery, but somewhere in between\cite{cattani2008core}. This is especially interesting as this aligns with our hypothesis that intermediate values of $Q$ are optimal.

\noindent\\\textit{\textbf{Teams can stagnate without new members.}}\\
Both interview participants indicated that new members were essential in bringing new ideas to the team. A highly static team (or, a team with a high $Q$ value), runs the risk of falling behind the technology trends.

\noindent\\\textit{\textbf{Collaboration is key to performance.}}\\
Both interviewees suggested that team collaboration had a direct impact on performance. Teams that showed good collaboration both intra-team and inter-team were noted for their positive performance characteristics. Further exploration of this topic is needed to determine more fully the nature of this collaboration. Specifically, we know that high conflict is detrimental to team performance, especially in complex work environments\cite{de2003task}. A further research direction may be whether or not good team collaboration/communication promotes conflict avoidance, conflict resolution, or otherwise.

\noindent\\\textit{\textbf{Priority, focus, and timeline are strong drivers.}}\\
Focus was specifically called out as an indicator of performance in one interview. Clear goals and a common objective were positive influencer, while continuously shifting focus was a negative influence. Research in this area suggests that creative teams have ``goals that are clear and compelling, but also open and challenging''\cite{isaksen2002climate}.

Timeline and priority discussed as strong contributors to team success. Team building could be thought of as function of the timeline, meaning that as the timeline of the project increased, more time could be devoted to team building activities.

It was also indicated that priority was a major factor. Different project priorities, and the resultant shifting of resources, had the possible effect of the weakening of string team's core members. This can be related to focus as well, as changing priorities can cause resource realignment and decreased overall focus.

\section{Conclusion}
Our hypothesis theorizes that medium values of clustering between collaborators on projects is optimal for project performance. The data analysis, specifically figures \ref{fig:q_fc_graph} and \ref{fig:q_sf_graph}, seem to support this claim. Also, qualitative research seems to support the hypothesis as well. Our interviews revealed that from experience, teams perform best when a stable core is periodically augmented with new members. This would be a qualitative representation of a intermediate $Q$ value -- significant clustering but not to the extent of isolation. Given both the quantitative and qualitative analysis, we can assert to align with Uzzi's findings\cite{uzzi2005collaboration} and show that intermediate values of $Q$ are correlated with increased project performance.

\section{Future Research}
Social networks, interpersonal relationships, and software development are all complex topics. We have explored a narrow slice of these in context, and there are some important areas of future exploration. First, the contributor graph contains a large amount of subgraphs, and as a whole appears less clustered than some other observed social network graphs -- consistent with Madey's findings\cite{madey2002open} on SourceForge data. As noted in our analysis, a large number of these subgraphs have only one or two contributors. This implies that a large number of contributors work alone and on a single project. We do not, with the current analysis, know how (or if) this degrades the quality of our findings.\\

A further area of study would be how our findings relate to previously defined \textit{success factors} in projects\cite{cooke2002real}. Also, we do not know directly \textit{how} intermediate values of $Q$ influence performance -- we can only observe a quantitative and qualitative correlation. We make a tacit equivalence to clustering and knowledge sharing, but we have not directly observed that in our analysis. Research does suggest that social factors are knowledge sharing motivators\cite{hendriks1999share}, but how that relates specifically to our work is unexplored here.\\

\section{Acknowledgments}
We thank Praveen Mittal and John Kaman for their insight, guidance, and instruction, M Iftekhar (Ifte) Rana and Traci St. Martin for their time and expertise, and Rick Kiefer for his review.

\bibliographystyle{plain}
\bibliography{bibliography}

\input{interview1}
\input{interview2}

\end{document}